---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file 
To regenerate, run `quarto::quarto_render("Readme.Rmd", output_file = "README.md")`
-->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rabbit

<!-- badges: start -->
<!-- badges: end -->

The package `rabbit` provides functions to standardise raw accelerometer data and extract movement dynamics from it. Accelerometer data is often collected in a raw format that may not be immediately suitable for analysis. The `standardize_data` function reads in raw accelerometer data, standardises the column names, and converts the timestamp to a consistent format. The `extract_movement_dynamics` function takes the standardised data and calculates various movement dynamics metrics, such as mean acceleration, variance, covariance, Overall Dynamic Body Acceleration (ODBA), and Vectorial Dynamic Body Acceleration (VDBA) using rolling window calculations. The data is then returned in a tidy format, ready for further analysis or classification of behaviour.

Accelerometer data is often collected at high frequencies, resulting in large datasets. The `rabbit` package is designed to efficiently process these large datasets by leveraging the `RcppRoll` package for optimized rolling window calculations. This allows users to quickly extract meaningful movement dynamics metrics from their accelerometer data, even when dealing with extensive recordings.


## Installation

You can install the development version of `rabbit` from GitHub using the `remotes` package:

```{r install, eval=FALSE}
# install.packages("remotes")
remotes::install_github("traitecoevo/rabbit")
```

## Example

This is a basic example file is less than one hour of a bilby called piccolo:


```{r example}
library(rabbit)
library(dplyr)

file_in = system.file("extdata", "raw_Pic2Jan_50000.parquet", package = "rabbit")

df <- 
  standardize_data(file_in = file_in, vars = c("Timestamp","accX","accY","accZ")) |> 
  extract_movement_dynamics()

nrow(df)

df |> filter(!is.na(time))
```

Alternatively, you can read in the data and then standardise it:

```{r}
library(arrow)
df <- arrow::read_parquet(file_in)
df_std <- standardize_data(df, vars = c("Timestamp","accX","accY","accZ"))
df_mvt <- extract_movement_dynamics(df_std)

```

## Identifying high sumVDBA times

sumVDBA is the best measure we have of heat-generating movement or activities:

```{r,warning=FALSE}
library(ggplot2)
df |> filter(!is.na(time)) |>
  ggplot(aes(x = time, y = sumVDBA)) +
  geom_point(size = 0.2) + theme_classic()
```

## Classifying behaviour

Data is now ready for classification of behaviour using your own behavioural classifier


