---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file 
To regenerate, run `quarto::quarto_render("Readme.Rmd", output_file = "README.md")`
-->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rabbit

<!-- badges: start -->
<!-- badges: end -->

The goal of rabbit is to ...

Use package `RcppRoll` to optimise claulcations of metrics clculated in wolling windows, e.g. rolling mean, var, sd etc.

## Installation

You can install the development version of rabbit from [GitHub](https://github.com/) with:

```{r install, eval=FALSE}
# install.packages("devtools")
 devtools::install_github("traitecoevo/rabbit")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(rabbit)
library(dplyr)

#Read file
data <- arrow::read_parquet("tests/testthat/raw_Pic2Jan_50000.parquet") |> slice(1:1000)

# Calcualte metrics
data_metrics <- moving_window_calcs_2(data)

data_metrics %>% slice(50:150)
```

Speed comparison on test file, comparing 3 techniques

```{r}
# source("tests/testthat/doAccloop.R")
# 
# microbenchmark::microbenchmark(
#   doAccloop_all(data),
#   moving_window_calcs(data),
#   moving_window_calcs_2(data)
# )
```


Speed comparison on a big test file, comparing the 2 faster techniques
```{r}

# df <- arrow::read_parquet('/Users/dfalster/GitHub/projects/Adams-Accelerometer_2024/data/Pic2Jan_S1.parquet')
# nrow(df)
```

version 1
```{r}
#system.time(moving_window_calcs(df))
```

version 2
```{r}
df <- arrow::read_parquet("tests/testthat/raw_Pic2Jan_50000.parquet")
system.time(dat <- moving_window_calcs_2(df))
```
## Identifying high sumODBA times
```{r,warning=FALSE}
library(ggplot2)
dat %>%
  ggplot(aes(x = time, y = sumVDBA)) +
  geom_point() + theme_classic()
```
```{r,warning=FALSE}
nighttime_activities <- classify_behaviors(dat, MSOM_path = "tests/testthat/MSOM_8by7.rda")

nighttime_activities <- nighttime_activities %>%
  mutate(behavior = forcats::fct_reorder(behavior, sumVDBA, .fun = median, na.rm = TRUE))

nighttime_activities %>%
  filter(!is.na(behavior)) %>%
  ggplot(aes(x = time, y = sumVDBA, col = behavior)) +
  geom_point(alpha = 0.7) + theme_classic()
```
