---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file 
To regenerate, run `quarto::quarto_render("Readme.Rmd", output_file = "README.md")`
-->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rabbit

<!-- badges: start -->
<!-- badges: end -->

The goal of rabbit is to ...

Use package `RcppRoll` to optimise claulcations of metrics clculated in wolling windows, e.g. rolling mean, var, sd etc.

## Installation

You can install the development version of rabbit from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("traitecoevo/rabbit")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(rabbit)
library(dplyr)

#Read file
data <- arrow::read_parquet("tests/testthat/raw_Pic2Jan_50000.parquet") |> slice(1:1000)

# Calcualte metrics
data_metrics <- moving_window_calcs_2(data)

data_metrics %>% slice(50:150)
```

Speed comparison on test file, comparing 3 techniques

```{r}
source("tests/testthat/doAccloop.R")

microbenchmark::microbenchmark(
  doAccloop_all(data),
  moving_window_calcs(data),
  moving_window_calcs_2(data)
)
```


Speed comparison on a big test file, comparing the 2 faster techniques
```{r}
df <- arrow::read_parquet('/Users/dfalster/GitHub/projects/Adams-Accelerometer_2024/data/Pic2Jan_S1.parquet')
nrow(df)
```

version 1
```{r}
system.time(moving_window_calcs(df))
```

version 2
```{r}
system.time(moving_window_calcs_2(df))
```
